---
title: "Assignment 5 Part 2"
format: html
editor: visual
---

## Setup

```{r}
library(tidyverse)
library(readxl)

df <- read_excel("c2015.xlsx")
```

## Question 1 - Missing SEX values already in NA form

```{r}
sum(is.na(df$SEX))
```

## Question 2 - Find other missing values in SEX and convert them to NA

```{r}
table(df$SEX, useNA = "ifany")

df <- df %>%
  mutate(
    SEX = na_if(SEX, "Unknown"),
    SEX = na_if(SEX, "Not Reported")
  )

table(df$SEX, useNA = "ifany")
```

## Question 3 - Replace missing SEX with majority sex

```{r}
table(df$SEX)

majority_sex <- names(which.max(table(df$SEX)))

majority_sex

df <- df %>%
  mutate(
    SEX = replace_na(SEX, majority_sex)
  )

table(df$SEX, useNA = "ifany")
```

## Question 4 - Examine AGE and convert missing forms to NA

```{r}
table(df$AGE, useNA = "ifany")

df <- df %>%
  mutate(
    AGE = na_if(AGE, "Unknown"),
    AGE = na_if(AGE, "Not Reported")
  )
```

## Question 5 - Change Less than 1 to "0"

```{r}
df <- df %>%
  mutate(
    AGE = str_replace(AGE, "Less than 1", "0")
  )

table(df$AGE, useNA = "ifany")
```

## Question 6 - Convert AGE to numeric

```{r}
class(df$AGE)

df <- df %>%
  mutate(
    AGE = as.numeric(AGE)
  )

class(df$AGE)
```


## Question 7 - Convert AGE to numeric

```{r}
class(df$AGE)

df <- df %>%
  mutate(
    AGE = as.numeric(AGE)
  )

class(df$AGE)

```

## Question 8 - Convert AGE to numeric

```{r}
mean_age <- mean(df$AGE, na.rm = TRUE)

mean_age

df <- df %>%
  mutate(
    AGE = replace_na(AGE, mean_age)
  )

sum(is.na(df$AGE))
```


## Question 9 - Convert AGE to numeric

```{r}
table(df$TRAV_SP, useNA = "ifany")

df <- df %>%
  mutate(
    TRAV_SP = str_remove(TRAV_SP, " MPH")
  )

table(df$TRAV_SP, useNA = "ifany")

df <- df %>%
  filter(TRAV_SP != "Greater" | is.na(TRAV_SP))

df <- df %>%
  filter(!str_detect(TRAV_SP, "Greater") | is.na(TRAV_SP))

df <- df %>%
  mutate(
    TRAV_SP = str_replace(TRAV_SP, "Stopped", "0")
  )

table(df$TRAV_SP, useNA = "ifany")

df <- df %>%
  mutate(
    TRAV_SP = na_if(TRAV_SP, "Unknown"),
    TRAV_SP = na_if(TRAV_SP, "Not Reported")
  )

class(df$TRAV_SP)

df <- df %>%
  mutate(
    TRAV_SP = as.numeric(TRAV_SP)
  )

class(df$TRAV_SP)

median_speed <- median(df$TRAV_SP, na.rm = TRUE)

median_speed

df <- df %>%
  mutate(
    TRAV_SP = replace_na(TRAV_SP, median_speed)
  )

sum(is.na(df$TRAV_SP))
```

## Question 10 - Convert AGE to numeric

```{r}
drivers <- df %>%
  filter(SEAT_POS == "Front Seat, Left Side") %>%
  mutate(
    AGE = as.numeric(AGE),
    TRAV_SP = as.numeric(TRAV_SP)
  )

cor(
  drivers$AGE,
  drivers$TRAV_SP,
  use = "complete.obs"
)

```
