---
title: "Assignment 3"
format: html
editor: visual
---

***How to do it?***:

-   Open the Quarto file of this assignment ([link](assignment3.qmd)) in Rstudio.

-   Right under each **question**, insert a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question.

-   `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.

-   Publish the html file to your Githiub Page.

***Submission***: Submit the link on Github of the assignment to Canvas.

```{r setup}
knitr::opts_chunk$set(message = FALSE)
```

------------------------------------------------------------------------

## Problems

\

1.  Create the following data frame

| Rank | Age | Name  |
|:-----|:----|:------|
| 0    | 28  | Tom   |
| 1    | 34  | Jack  |
| 2    | 29  | Steve |
| 3    | 42  | Ricky |

```{r}
## ---- 1. Create the data frame ----
df1 <- data.frame(
  Rank = 0:3,
  Age  = c(28, 34, 29, 42),
  Name = c("Tom", "Jack", "Steve", "Ricky")
)
df1
```

2.  Use `read.csv` to import the Covid19 Vaccination data from WHO: [link](https://bryantstats.github.io/math421/data/us-states.csv).

```{r}
df <- read.csv('https://bryantstats.github.io/math421/data/us-states.csv')
```

3.  Show the names of the variables in the data

```{r}

names(df)
```

4.  How many columns and rows the data have?

```{r}
nrow(df)
ncol(df)
dim(df)
```

5.  How many missing values are there? Show the missing values by columns. What variable has the most number of missing values?

```{r}
sum(is.na(df))
colSums(is.na(df))
names(which.max(colSums(is.na(df))))
```

6.  What is the class of the `date` column. Change the `date` columns to `date` type using the `as.Date` function. Show the new class of the `date` column.

```{r}
class(df$date)
df$date <- as.Date(df$date)
class(df$date)
```

7.  Capitalize the names of all the variables

```{r}
df_caps <- df
names(df_caps) <- toupper(names(df_caps))
names(df_caps)
```

8.  Find the average number of cases per day. Find the maximum cases a day.

```{r}
mean(df$cases)
max(df$cases)
```

9.  How many states are there in the data?

```{r}
length(unique(df$state))
```

10. Create a new variable `weekdays` to store the weekday for each rows.

```{r}
df$weekdays <- weekdays(df$date)
head(df$weekdays)
```

11. Create the categorical variable `death2` variable taking the values as follows

-   `has_death` if there is a death that day
-   `no_death` if there is no death that day

Find the frequency and relative frequency of `no_death` and `has_death`.

```{r}
df$death2 <- ifelse(df$deaths > 0, "has_death", "no_death")
table(df$death2)
prop.table(table(df$death2))
```

12. Find the first quartile (Q1), second quartile (Q2) and and third quartile (Q3) of the variable `death`. (Hint: Use the `summary` function)

```{r}
summary(df$deaths)
quantile(df$deaths, probs = c(0.25, 0.5, 0.75), na.rm = TRUE)
```

13. Create the categorical variable `death3` variable taking the values as follows

-   `low_death` if the number of deaths smaller than the 25 percentile (Q1)

-   `mid_death` if the number of deaths from Q1 to Q3

-   `high_death` if the number of deaths greater than Q3

```{r}
library(dplyr)

q <- quantile(df$deaths, probs = c(0.25, 0.75), na.rm = TRUE)

df$death3 <- case_when(
  df$deaths < q[1] ~ "low_death",
  df$deaths >= q[1] & df$deaths <= q[2] ~ "mid_death",
  df$deaths > q[2] ~ "high_death"
)
table(df$death3)
```

14. Find the average cases in Rhode Island in 2021

```{r}
ri_2021 <- df[df$state == "Rhode Island" & format(df$date, "%Y") == "2021", ]
mean(ri_2021$cases)
```

15. Find the median cases by weekdays in Rhode Island in 2021

```{r}
aggregate(cases ~ weekdays, data = ri_2021, FUN = median)
```

16. Compare the median cases in Rhode Island in June, July, August and September in 2021.

```{r}
ri_2021$month <- format(ri_2021$date, "%B")

ri_summer <- ri_2021[ri_2021$month %in% c("June", "July", "August", "September"), ]
aggregate(cases ~ month, data = ri_summer, FUN = median)
```

17. Find your own dataset, import it and implement the following functions on the data

-   head
-   str
-   names
-   mean, min, max
-   table
-   is.na
-   colSums
-   class
-   cor
-   by
-   ifelse
-   case_when

```{r}
titanic <- read.csv('https://bryantstats.github.io/math421/data/titanic.csv')

head(titanic)
str(titanic)
names(titanic)

mean(titanic$Age, na.rm = TRUE)
min(titanic$Age, na.rm = TRUE)
max(titanic$Age, na.rm = TRUE)

table(titanic$Sex)

is.na(titanic$Age)
colSums(is.na(titanic))

class(titanic$Age)

cor(titanic$Age, titanic$Fare, use = "complete.obs")

by(titanic$Fare, titanic$Pclass, mean)

titanic$AgeGroup <- ifelse(titanic$Age >= 18, "Adult", "Minor")

titanic$FareLevel <- case_when(
  titanic$Fare < 10 ~ "Low",
  titanic$Fare < 30 ~ "Medium",
  TRUE ~ "High"
)
```

18. In the dataset in #17, practice the follows. You can reuse the code of 17.

-   Create a categorical variable from a continuous variable
-   From one categorical variable, create a new categorical variable with fewer categories

```{r}
# continuous -> categorical
titanic$AgeGroup <- ifelse(titanic$Age >= 18, "Adult", "Minor")
table(titanic$AgeGroup)

# categorical -> fewer categories
titanic$ClassGroup <- case_when(
  titanic$Pclass == 1 ~ "Upper",
  titanic$Pclass %in% c(2, 3) ~ "Lower/Middle"
)
table(titanic$ClassGroup)
```

If you do not have a data, you can use [the titanic dataset](https://www.kaggle.com/competitions/titanic/overview), which can be downloaded at [this link](https://bryantstats.github.io/math421/data/titanic.csv)

```{r}
df = read.csv('https://bryantstats.github.io/math421/data/us-states.csv')
names(df)

str(df1)

names(df1)

mean(df1$Age)

table(df1$Sex)

table(df1$Survived)

table(df1$Pclass)



```
