---
title: "Assignment 4 Extra Credit"
format: html
editor: visual
---

```{r}
library(tidyverse)

df <- read_csv("imdb_top_1000.csv")

df <- df %>%
  mutate(
    Gross = as.numeric(str_remove_all(Gross, ",")),
    Runtime = as.numeric(str_remove(Runtime, " min")),
    Released_Year = as.numeric(Released_Year)
  )
```

## Question 1


```{r}
names(df)
```


## Question 2


```{r}
df %>%
  select(Series_Title, Gross) %>%
  arrange(desc(Gross)) %>%
  head(10)
```


## Question 3


```{r}
min(df$IMDB_Rating, na.rm = TRUE)

df %>%
  filter(IMDB_Rating == min(IMDB_Rating, na.rm = TRUE)) %>%
  select(Series_Title, IMDB_Rating) %>%
  head(5)
```

## Question 4 Which year have the most number of movies released in the list? What is the total of money earned on that year?


```{r}
df %>%
  group_by(Released_Year) %>%
  summarize(
    SeriesTitle = n(),
    Total_Gross = sum(Gross,na.rm = TRUE),
  )%>%
  arrange(desc(SeriesTitle)) %>%
  head(1)

```


## Question 5 - Average gross earnings per movie


```{r}
df %>%
  group_by(Series_Title) %>%
  summarize(
    Average_Gross = mean(Gross, na.rm = TRUE)
  )
```

## Question 6 - Calculate the average number of votes by year. Calculate the average number of votes of movies that have IMDB rating greater than 9.


```{r}
df %>%
  group_by(Released_Year) %>%
  summarize(
    Avg_votes = mean(No_of_Votes,na.rd = TRUE)
  ) 
  
df %>%
   filter(IMDB_Rating > 9) %>%
   summarize(
     Avg_votes2 = mean(No_of_Votes, na.rm = TRUE)
   )
```


## Question 7 -Calculate the average Meta score in 2020 of movies that have number of votes in the third quartile.


```{r}
Q2 = quantile(df$No_of_Votes, 0.50)
Q3 = quantile(df$No_of_Votes, 0.75)

df %>%
  filter(
    Released_Year == 2020,
    
  )%>%


```


## Question 8 - (Optional - Challenging). The current Runtime variable is not a numeric. Use the str_remove function to remove min from the variables then use as.numeric to convert the variable to numeric. Calculate the average running time in the 2010s. Calculate the correlation between running time and rating (adding use="complete.obs" in the cor function to ignore the missing values).


```{r}

```



## Question 9 - We can use select_if to select columns satisfying a condition and use summarise_if to do calculation on columns satisfying a condition. Try the follows to understand these functions.


```{r}

```


## Question 10 - Implement the follows functions or combos. Drawing a comment or summary from each calculation. The codes in this question should be different from the codes used in other questions.


```{r}

```






