
---
title: "Assignment 4"
format: html
editor: visual
---

**How to do it?**

- Open this Quarto file in RStudio.
- Run the code chunks below.
- Click **Render** to export the assignment as HTML.
- Publish the HTML file to your GitHub Pages site.

**Submission:** Submit the GitHub Pages link to Canvas.

```{r setup}
#| message: false
#| warning: false

library(tidyverse)
library(lubridate)

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
```

---

#### 1. Install `tidyverse` package

Install `tidyverse` if it is not already installed.

```{r}
if (!requireNamespace("tidyverse", quietly = TRUE)) {
  install.packages("tidyverse")
}

library(tidyverse)
```

---

#### 2. Read the data using `read_csv`

Import the US COVID-19 dataset.

```{r}
df <- read_csv(
  "https://bryantstats.github.io/math421/data/all-states-history.csv",
  show_col_types = FALSE
)

head(df)
```

---

#### 3. Fix the date and create new variables

Create `month`, `weekday`, and `monthday`.

```{r}
library(lubridate)

df$date <- ymd(df$date)

# Month of the year
df$month <- month(df$date)

# Day of the week
# Sunday = 1, Saturday = 7
df$weekday <- wday(df$date)

# Day of the month
df$monthday <- mday(df$date)

head(df)
```

---

#### 4. Create new variables with `case_when`

**Create the daily_death variable.**

Classify daily deaths as low, medium, or high.

```{r}
library(dplyr)

df$daily_death <- case_when(
  is.na(df$deathIncrease) ~ NA_character_,
  df$deathIncrease < 3 ~ "low",
  df$deathIncrease <= 14 ~ "medium",
  TRUE ~ "high"
)

table(df$daily_death)
```

**Create the month2 variable.**

- early_month: days 1–10
- mid_month: days 11–20
- end_month: days 21–31

```{r}
df$month2 <- case_when(
  df$monthday <= 10 ~ "early_month",
  df$monthday <= 20 ~ "mid_month",
  TRUE ~ "end_month"
)

table(df$month2)
```

**Create the weekend variable.**

Assign 1 to Saturday and Sunday and 0 to other days.

```{r}
df$weekend <- case_when(
  df$weekday %in% c(1, 7) ~ 1,
  TRUE ~ 0
)

table(df$weekend)
```

---

#### 5. Select function

Remove the `totalTestsViral` column.

```{r}
df <- df %>%
  select(-totalTestsViral)

head(df)
```

---

#### 6. Pipe Operator (%>%)

Rewrite the calculations using the pipe operator.

```{r}
x <- 1:100

# Sum
x %>% sum()

# Square root
x %>% sqrt()

# Square root followed by sum
x %>%
  sqrt() %>%
  sum()

# Square root, sum, and logarithm
x %>%
  sqrt() %>%
  sum() %>%
  log()

# Logarithm of 16 with base 2
16 %>% log(2)
```

---

#### 7. Combo 1: group_by + summarise

**Calculate the median number of cases by month.**

```{r}
df %>%
  group_by(month) %>%
  summarise(
    median_cases = median(
      positiveIncrease,
      na.rm = TRUE
    )
  )
```

**Calculate the average number of cases by month2.**

```{r}
df %>%
  group_by(month2) %>%
  summarise(
    avg_cases = mean(
      positiveIncrease,
      na.rm = TRUE
    )
  )
```

**Calculate the median number of cases by weekend.**

```{r}
df %>%
  group_by(weekend) %>%
  summarise(
    median_cases = median(
      positiveIncrease,
      na.rm = TRUE
    )
  )
```

---

#### 8. Combo 2: filter + group_by + summarise

**Calculate the median number of cases on weekends in October and November 2020, grouped by month.**

```{r}
df %>%
  filter(
    year(date) == 2020,
    month %in% c(10, 11),
    weekend == 1
  ) %>%
  group_by(month) %>%
  summarise(
    median_cases = median(
      positiveIncrease,
      na.rm = TRUE
    )
  )
```

**Calculate the average number of deaths during different periods of a month in Fall 2020.**

```{r}
df %>%
  filter(
    year(date) == 2020,
    month %in% 9:11
  ) %>%
  group_by(month2) %>%
  summarise(
    avg_death = mean(
      deathIncrease,
      na.rm = TRUE
    )
  )
```

**Compare average hospitalizations between weekdays and weekends in Summer 2020.**

```{r}
df %>%
  filter(
    year(date) == 2020,
    month %in% 6:8
  ) %>%
  group_by(weekend) %>%
  summarise(
    avg_hosp = mean(
      hospitalizedCurrently,
      na.rm = TRUE
    )
  )
```

**Redo Questions 14 and 15 from Assignment 3.**

Import the dataset used in Assignment 3.

```{r}
df3 <- read_csv(
  "https://bryantstats.github.io/math421/data/us-states.csv",
  show_col_types = FALSE
)

df3$date <- as.Date(df3$date)

df3$weekdays <- weekdays(df3$date)
```

**Question 14: Calculate the average number of cases in Rhode Island in 2021.**

```{r}
df3 %>%
  filter(
    state == "Rhode Island",
    year(date) == 2021
  ) %>%
  summarise(
    avg_cases = mean(
      cases,
      na.rm = TRUE
    )
  )
```

**Question 15: Calculate the median number of cases by weekday in Rhode Island in 2021.**

```{r}
df3 %>%
  filter(
    state == "Rhode Island",
    year(date) == 2021
  ) %>%
  group_by(weekdays) %>%
  summarise(
    median_cases = median(
      cases,
      na.rm = TRUE
    )
  )
```

---

#### 9. Combo 3: filter + group_by + summarise + arrange

Find the month with the highest number of deaths on weekends.

```{r}
df %>%
  filter(weekend == 1) %>%
  group_by(month) %>%
  summarise(
    total_death = sum(
      deathIncrease,
      na.rm = TRUE
    )
  ) %>%
  arrange(desc(total_death))
```

---

#### 10. Use your own dataset

For this question, we will use the Titanic dataset to demonstrate all the required functions and combinations.

**Import the dataset.**

```{r}
t <- read_csv(
  "https://bryantstats.github.io/math421/data/titanic.csv",
  show_col_types = FALSE
)

head(t)
```

**1. Select**

Select survival status, passenger class, sex, age, and fare.

```{r}
t %>%
  select(Survived, Pclass, Sex, Age, Fare)
```

**2. Filter**

Select female passengers in first class.

```{r}
t %>%
  filter(
    Sex == "female",
    Pclass == 1
  )
```

**3. Mutate**

Create a variable representing total family size.

```{r}
t <- t %>%
  mutate(
    family_size = SibSp + Parch + 1
  )

head(t)
```

**4. Summarise**

Calculate average age and average fare.

```{r}
t %>%
  summarise(
    avg_age = mean(Age, na.rm = TRUE),
    avg_fare = mean(Fare, na.rm = TRUE)
  )
```

**5. Arrange**

Sort passengers by fare in descending order.

```{r}
t %>%
  arrange(desc(Fare))
```

**6. Count**

Count the number of passengers in each class.

```{r}
t %>%
  count(Pclass)
```

**7. Count + Arrange**

Count passengers by embarkation port and sort by frequency.

```{r}
t %>%
  count(Embarked) %>%
  arrange(desc(n))
```

**8. Filter + Count + Arrange**

Count surviving passengers by class and sort in descending order.

```{r}
t %>%
  filter(Survived == 1) %>%
  count(Pclass) %>%
  arrange(desc(n))
```

**9. Group_by + Summarise**

Calculate survival rates by sex.

```{r}
t %>%
  group_by(Sex) %>%
  summarise(
    survival_rate = mean(
      Survived,
      na.rm = TRUE
    )
  )
```

**10. Filter + Group_by + Summarise**

Calculate average fares by passenger class for adults aged 18 and older.

```{r}
t %>%
  filter(Age >= 18) %>%
  group_by(Pclass) %>%
  summarise(
    avg_fare = mean(
      Fare,
      na.rm = TRUE
    )
  )
```

**11. Filter + Group_by + Summarise + Arrange**

Calculate survival rates by passenger class and sex, excluding passengers with missing ages.

```{r}
t %>%
  filter(!is.na(Age)) %>%
  group_by(Pclass, Sex) %>%
  summarise(
    survival_rate = mean(
      Survived,
      na.rm = TRUE
    ),
    .groups = "drop"
  ) %>%
  arrange(desc(survival_rate))
```
