
---
title: "Assignment 2"
format: html
editor: visual
---

***How to do it?***:

-   Open the Quarto file of this assignment ([link](assignment2q.qmd)) in Rstudio.

-   Right under each **question**, insert a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question.

-   `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.

-   Publish the html file to your Githiub Page.

***Submission***: Submit the link on Github of the assignment to Canvas.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

------------------------------------------------------------------------

## Problems

1.  Calculate:
```{r}
Prob1 = seq(1,100,3)
Prob1
sum(Prob1)
```



$$
S = 1 + 4 + 7 + 10 +... + 100
$$

2.  Calculate:

```{r}
Prob2 = seq(1,100,1)
Prob2^100
sum(Prob2^100)
```

$$
S = 1^{100} + 2^{100} + 3^{100} +...+100^{100}
$$

3.  Calculate:

```{r}
Prob3Even = seq(1,100,1)
Prob3Even
Prob3Odd = seq(2,101,1)
Prob3Odd
Prob3Rep = rep(1,50)
Prob3Rep
Prob3Rep/(Prob3Even*Prob3Odd)
sum(Prob3Rep/(Prob3Even*Prob3Odd))
```

$$
S = \frac{1}{1\cdot 2}+\frac{1}{2 \cdot 3}+...+\frac{1}{100\cdot 101}
$$

4.  Calculate:

```{r}
Prob40 = seq(1,100,1)
Prob40
Prob41 = seq(2,101,1)
Prob41
Prob42 = seq(3,102,1)
Prob42
Prob4Rep = rep(1,100)
Prob4Rep/(Prob40*Prob41*Prob42)
sum(Prob4Rep/(Prob40*Prob41*Prob42))
```

$$
S = \frac{1}{1\cdot 2 \cdot 3}+\frac{1}{2 \cdot 3 \cdot 4}+...+\frac{1}{100\cdot 101 \cdot 102}
$$

5.  Calculate:

```{r}
Prob5 = seq(1,101,2)
Prob5^Prob5
sum(Prob5^Prob5)
```

$$
S = 1^1 + 3^3 + 5^5 + 7^7 +....+101^{101}
$$