---
title: "Assignment 5 Part 1"
format: html
editor: visual
---

## Question 1 - Plot the number of NAs using aggr

```{r}

library(tidyverse)
library(VIM)

df <- read_csv("adult_census_missing.csv")

colSums(is.na(df))
```

## Question 2 - Plot the number of NAs using aggr

```{r}
aggr(
  df,
  numbers = TRUE
)
```

## Question 3 Find other forms of missing values. Hint: You can use the table function to check if there are suspicious categories (Unknown, for example) in a variable. What all other forms of missing values shown in the data.

```{r}
df_median_majority %>%
  select_if(is.character) %>%
  map(table)

df %>%
  summarise(
    across(
      where(is.character),
      ~ sum(.x == "?", na.rm = TRUE)
    )
  )
```

## Question 4 Replace all the forms of missing values found to NA

```{r}
df <- df %>%
  mutate(
    across(
      where(is.character),
      ~ na_if(.x, "?")
    )
  ) %>%
  mutate(
    across(
      where(is.character),
      ~ na_if(.x, "Unknown")
    )
  ) %>%
  mutate(
    across(
      where(is.character),
      ~ na_if(.x, "")
    )
  )
colSums(is.na(df))
```

## Question 5 Replot the number of NAs for each column.

```{r}
aggr(
  df,
  numbers = TRUE
)
```

## Question 6 Approach 1 to handle NAs: remove all rows that have any NAs. Save the dataset after removing as a different data. The original data is unchanged (still have NAs). How many rows left after removing?

```{r}
df_removed <- df %>%
  drop_na()

nrow(df_removed)

nrow(df)

nrow(df_removed)
```

## Question 7 Approach 2 to handle NAs: Fill all the NAs with the previous or next value. (Hint: Using fill function). Save the dataset after filling as a different data. The original data is unchanged (still have NAs).

```{r}
df_filled <- df %>%
  fill(
    everything(),
    .direction = "downup"
  )
colSums(is.na(df_filled))
```

## Question 8 Approach 3 to handle NAs: For numeric variable, replace the NAs by the median. For categorical variables, replace the NAs to the majority.

```{r}
df_median_majority <- df

Unknown = "Unknown0"

df_median_majority <- df_median_majority %>%
  mutate(
    across(
      where(is.character),
      ~ replace_na(.x, Unknown)
    )
  )

get_majority <- function(x) {
  names(sort(table(x), decreasing = TRUE))[1]
}

##majority 1 is most common value
df_median_majority <- df_median_majority %>%
  mutate(
    across(
      where(is.character),
      ~ replace_na(.x, get_majority(.x))
    )
  )

colSums(is.na(df_median_majority))
```
